Windowed least-squares model reduction for dynamical systems

### 2020-08-25 numerical_experiments.tex
minor language changes
working on revision

\subsubsection{Numerical results}
We first assess the impact of the window size on the performance of the \methodAcronymROMs.  We consider a set of \methodAcronymROMs\ that minimize the residual over windows of constant size 
$\DeltaSlabArg{n} \equiv \DeltaSlabArg{} =  .002$, $0.004$ ,$0.008$, $0.02$, $0.04$, $0.10$, $0.20$, $1.0$. We additionally consider the standard Galerkin and LSPG ROMs. We first show results for \methodAcronymROMs\ using the direct method with CN time discretization; a comparison of different time-marching methods and direct/indirect solution techniques will be provided later in this section. 
First, Figure~\ref{fig:sod_density} presents the density solutions produced by the various ROMs at $t = 0.5$ and $1.0$. Figure~\ref{fig:sod_xt} shows 
$x-t$ diagrams for the same density solutions. From Figures~\ref{fig:sod_density} and~\ref{fig:sod_xt}, we observe that the LSPG and Galerkin ROMs accurately characterize 
the system: they correctly track the shock location, expansion waves, etc. We observe both predictions, however, to be highly oscillatory. These oscillations are 
not physical and can lead to numerical instabilities; e.g., due to negative pressure. We observe the \methodAcronymROMs\ to produce less oscillatory solutions than both 
the Galerkin and LSPG ROMs. Critically, we see that the solution becomes less oscillatory as the window size over which the residual is minimized grows. The solution displays no oscillations when the residual is minimized over the entire space--time domain.

@@

Finally, we study the impact of the time step on the \methodAcronymROM\ results. We examine \methodAcronymROMs\ that use a
window size of $\Delta T = 0.1$, with time steps $\Delta t =$  $0.001$, $0.002$, $0.005$, $0.01$, $0.05$, $0.1$ (i.e.,  
$100$, $50$, $20$, $10$, $2$, and $1$ time instances per window). We additionally 
consider LSPG ROMs leveraging the same set of time steps. To assess time step convergence, we compare results to a new full-order model, which is as described in Section~\ref{sec:sod_fom} but uses a fine time step of $\Delta t = 10^{-4}$. Figure~\ref{fig:time_step_study} shows the $\elltwo$-error and residual norm  
of the various ROMs. We observe that the $\elltwo$-error and residual norm 
of the \methodAcronymROMs\ decrease and converge as the time step decreases. This is in direct contrast to the LSPG ROMs, in where the $\elltwo$-error and residual norm display a complex dependence on the time step. This result 
demonstrates that \methodAcronym\ overcomes the time-step dependence inherit to the LSPG approach. Lastly, Figure~\ref{fig:walltime_dtvar}
shows the relative wall-clock times of the \methodAcronymROMs\ with respect to the LSPG ROMs. It is seen that for all time steps considered, the \methodAcronymROMs\ 
are less than 6x the cost of LSPG; this is despite the window sizes comprising up to 100 time instances. 
\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/error_vs_window_dtvar.pdf}
\caption{Space--time $\elltwo$-error of the \methodAcronymROM.}
\label{fig:sod_error_methods_a}
\end{subfigure}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/objective_vs_window_dtvar.pdf}
\caption{Space--time residual norm of the \methodAcronymROM.} 
\label{fig:sod_error_methods_b}
\end{subfigure}
\caption{Performance metrics of the Galerkin, LSPG, and \methodAcronymROMs.} 
\label{fig:time_step_study}
\end{center}
\end{figure}

@@

>>>>

\subsubsection{Numerical results}
We first assess the impact of the window size on the performance of the \methodAcronymROMs.  We consider a set of \methodAcronymROMs\ that minimize the residual over windows of constant size 
$\DeltaSlabArg{n} \equiv \DeltaSlabArg{} =  .002$, $0.004$ ,$0.008$, $0.02$, $0.04$, $0.10$, $0.20$, $1.0$. We additionally consider the standard Galerkin and LSPG ROMs. We first show results for \methodAcronymROMs\ using the direct method with CN time discretization; a comparison of different time-marching methods and direct/indirect solution techniques will be provided later in this section. 
First, Figure~\ref{fig:sod_density} presents the density solutions produced by the various ROMs at $t = 0.5$ and $1.0$. Figure~\ref{fig:sod_xt} shows 
$\mathsf{x}-t$ diagrams for the same density solutions. From Figures~\ref{fig:sod_density} and~\ref{fig:sod_xt}, we observe that the LSPG and Galerkin ROMs accurately characterize 
the system: they correctly track the shock location, expansion waves, etc. We observe both predictions, however, to be highly oscillatory. These oscillations are 
not physical and can lead to numerical instabilities; e.g., due to negative pressure. We observe the \methodAcronymROMs\ to produce less oscillatory solutions than both 
the Galerkin and LSPG ROMs. Critically, we see that the solution becomes less oscillatory as the window size over which the residual is minimized grows. The solution displays no oscillations when the residual is minimized over the entire space--time domain.

@@

Next, we study the impact of the time step on the \methodAcronymROM\ results. We examine \methodAcronymROMs\ that use a
window size of $\Delta T = 0.1$, with time steps $\Delta t =$  $0.001$, $0.002$, $0.005$, $0.01$, $0.05$, $0.1$ (i.e.,  
$100$, $50$, $20$, $10$, $2$, and $1$ time instances per window). We additionally 
consider LSPG ROMs leveraging the same set of time steps. To assess time step convergence, we compare results to a new full-order model, which is as described in Section~\ref{sec:sod_fom} but uses a fine time step of $\Delta t = 10^{-4}$. Figure~\ref{fig:time_step_study} shows the $\elltwo$-error and residual norm  
of the various ROMs. We observe that the $\elltwo$-error and residual norm 
of the \methodAcronymROMs\ decrease and converge as the time step decreases. This is in direct contrast to the LSPG ROMs, in where the $\elltwo$-error and residual norm display a complex dependence on the time step. This result 
demonstrates that \methodAcronym\ overcomes the time-step dependence inherit to the LSPG approach. Lastly, Figure~\ref{fig:walltime_dtvar}
shows the relative wall-clock times of the \methodAcronymROMs\ with respect to the LSPG ROMs. It is seen that for all time steps considered, the \methodAcronymROMs\ 
are less than 6x the cost of LSPG; this is despite the window sizes comprising up to 100 time instances. 
\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/error_vs_window_dtvar.pdf}
\caption{Space--time $\elltwo$-error of the \methodAcronymROM.}
\label{fig:sod_error_methods_a}
\end{subfigure}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/objective_vs_window_dtvar.pdf}
\caption{Space--time residual norm of the \methodAcronymROM.} 
\label{fig:sod_error_methods_b}
\end{subfigure}
\caption{Performance metrics of the Galerkin, LSPG, and \methodAcronymROMs.} 
\label{fig:time_step_study}
\end{center}
\end{figure}

@@

\subsubsection{Results as a function of basis dimension}
Lastly, we assess the impact of the basis dimension on the performance of the various reduced-order models. We again consider S-reduction trial subspaces. We obtain the trial subspaces by evolving the FOM for $t \in [0.0,1.0]$ at a time-step of $\Delta t = 0.002$ with the CN scheme. We execute Algorithm~\ref{alg:pod} with inputs $N_{\text{skip}} = 2$, $\stateIntercept = \state_0$\footnote{We observed negative pressures in the projected initial conditions for trial subspaces of dimension $K=20,30,40$ if we used the offset $\stateIntercept = \bz$. Thus, for our convergence study we employed $\stateIntercept = \state_0$ so-as to avoid negative pressures in the projected initial conditions.}, and $\romdim = 10,20,30,40,50$. We investigate reduced-order models based on Galerkin projection, LSPG projection, and the WLS approach. All methods leverage the CN time scheme with $\Delta t = 0.002$. We present results for windows of size $\Delta T^n \equiv \Delta T = 0.002,0.004,0.02,0.04,0.10$ for the WLS approach. All WLS ROMs are solved via the direct approach described previously.

Figure~\ref{fig:convergence_study} shows the integrated $\ell^2$-error of the various ROM solutions, as well as the optimal solution as measured in the $\ell^2$-norm. In Figure~\ref{fig:sod_error_converge} we first observe that for a basis of dimension $\romdim = 10$, the Galerkin, LSPG, and WLS $\Delta T = 0.002$, $\Delta T = 0.004$ ROMs do not converge. Upon investigation, this lack of convergence occurred due to the presence of negative pressure. We next observe that the $\ell^2$-error monotonically decreases as the basis dimension increases for all remaining ROM solutions with the exception of the Galerkin, LSPG, and WLS $\Delta T = 0.002$, $\Delta T = 0.004$ at dimension $\romdim = 30$. Overall, we observe that larger window sizes yield lower errors when the basis dimension is small. For larger basis dimensions we observe all ROMs to yield accurate solutions, and a larger window size does not always yield a lower error. In Figure~\ref{fig:sod_residual_converge}, we next observe that the residual monotonically decreases as the basis dimension grows. This result is supported, although not guaranteed, by the error analysis presented in Section~\ref{sec:analysis} and Ref.~\cite{carlberg_lspg_v_galerkin}. Finally, we observe that larger window sizes yields lower integrated residuals.

\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/error_converge.pdf}
\caption{Space--time $\elltwo$-error of the \methodAcronymROM.}
\label{fig:sod_error_converge}
\end{subfigure}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/residual_converge.pdf}
\caption{Space--time residual norm of the \methodAcronymROM.} 
\label{fig:sod_residual_converge}
\end{subfigure}
\caption{Performance metrics of the Galerkin, LSPG, and \methodAcronymROMs\ as a function of basis dimension. Missing entries indicate that the solver did not converge.} 
\label{fig:convergence_study}
\end{center}
\end{figure}
